---
title: "simpleRegress"
author: "Robert Solis"
date: "7/5/23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
the bears neck is 20 inches around which would mean its about 57 in length. The secondary plot shows the residual standard error is 5.376 and The R Squared is .7524. The neck residual is the second best and the  r squared is also the second best.


## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
MYCHEST <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(MYCHEST)
summary.lm(MYCHEST)
```
 the residual standard error is 4.939 and is lower than the previous model which proeved to be better. Secondly the R squared is 0.791, which is higher then the previous result model .7524. By ranking we can conclude chest is higher than neck.For the residual error chest is best.
 
 
```{r}
MYHEADWTH <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(MYHEADWTH)
summary.lm(MYHEADWTH)
```

The third model residual standard is 7.32 which is worse. as well the r Squared is 0.5409 and is declining from the previous models. Overall this is the worst data chart so far. This model is the worst residual data and squared.

```{r}
MYWEIGHT <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(MYWEIGHT)
summary.lm(MYWEIGHT)
```
 This model shows the bears weight. the residual standard is  5.432, and the adjusted R-Square is 0.7472. Out of all the models ran chest,neck, headwth and weight. Model four has the third best stats in residual data and r squared.
 
```{r}
MYAGE <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(MYAGE)
summary.lm(MYAGE)
```
This model shows the bears age by length the residual standard is  7.511, and the adjusted R-Square is 0.5166.  This model rank is the fourth best out of all.
 
```{r}
MYHEADLEN <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(MYHEADLEN)
summary.lm(MYHEADLEN)
```
This models residual is 4.235 and the R squared is .08463. This ranking is the highest in standard error and highest in r squared results.


#conclsuions 
After taken into account all the new results
 From a highest to lowest rating in Standard error, model 6 is the highest by 4.235 then model 2 by 4.939, followed by model 1 by 5.376, then 4 by 5.432. Lastly  the second lowest model is 3 by 7.32 and the lowest is model 5 by 7.511
 
 The after taking into acount all the information gathered the Highest to lowest ratings for R squared is model 6 by 0.8463,	then model 2 by 0.791,	followed by model 1  by  0.7524, then model 4 by 7472. lastly  model 3 by 0.5409 and last model 5 by 0.5166.
 