---
title: "simpleRegress"
author: "Robert Solis"
date: "7/5/23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
the bears neck is 20 inches around which would mean its about 57 in length. The secondary plot shows the residual standard error is 5.376 and The R Squared is .7524. The neck residual is the second best and the  r squared is also the second best.


## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
MYCHEST <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(MYCHEST)
summary.lm(MYCHEST)
```
 the residual standard error is 4.939 and is lower than the previous model which proeved to be better. Secondly the R squared is 0.791, which is higher then the previous result model .7524. By ranking we can conclude chest is higher than neck.For the residual error chest is best.
 
 
```{r}
MYHEADWTH <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(MYHEADWTH)
summary.lm(MYHEADWTH)
```

The third model residual standard is 7.32 which is worse. as well the r Squared is 0.5409 and is declining from the previous models. Overall this is the worst data chart so far. This model is the worst residual data and squared.

```{r}
MYWEIGHT <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(MYWEIGHT)
summary.lm(MYWEIGHT)
```
 This model shows the bears weight. the residual standard is  5.432, and the adjusted R-Square is 0.7472. Out of all the models ran chest,neck, headwth and weight.
 
 
 